R Notebook

### 2021-02-04 examples/Toy-example-balanced-accessibility.Rmd
Added example of equity to the notebook.

The values check. Next, are there any hubs that serve zero population units? (same issue with division by zero):
```{r}
ttm_walk$sum_b2 %>% 
  summary()
```

@@

Check the balanced_impedance factors for the population units:
```{r}
ttm_walk %>% 
  dplyr::filter(hub == "46") 
```

@@

>>>>

The values check.

Check `sum_b1`:
```{r}
ttm_walk$sum_b1 %>% 
  summary()
```

Are there any hubs that serve zero population units? (same issue with division by zero):
```{r}
ttm_walk$sum_b2 %>% 
  summary()
```

@@

Check the balanced_impedance factors for the population units:
```{r}
ttm_walk %>% 
  dplyr::filter(hub == "46") %>%
  pull(balanced_impedance_2) %>%
  sum()
```

The sum should be 1.

@@

<!--The text that follows was mostly copied from PLoS ONE paper-->

## Disparity Analysis

An advantage of the use of adjusted weights for proportional allocation of demand and level of service it that, after rectifying the inflation artifact, they make it is possible to conduct accessibility disparity analysis in a very intuitive way.

For instance, an analyst interested in equity analysis could allocate the total level of service uniformly to every population unit. In other words, the total level of service (which equals the sum of accessibility over the system) can be divided by the number of population units in the system to return the Average Local Population Unit PPR. The resulting mean value, call it $L_i^e$ then would be assigned to the population units as their "equitable" share of the total level of service in the system. Next, the equitative distribution of the level of service in each population unit is substracted from the estimated accessibility to arrive at a disparity index. When the difference between these two quantities is positive, this would indicate that accessibility at that location exceeds its equitable share of level of service. On the other hand, when the difference is negative, accessibility at the location is below its equitable share of the level of service.

This approach is reminiscent of the Spatial Access Ratio (SPAR) proposed by Wan et al. [-@Wan2012SPAR], which is calculated as the ratio between a population center's accessibility and the mean accessibility across all population centers. Wan et al. [-@Wan2012] calculate SPAR based on the results of their 3SFCA method, by rescaling the accessibility measures to reflect the percentage difference in each population center's accessibility relative to the mean. This measure is designed to overcome the sensitivity of existing FCA metrics to the impedance function. In contrast, the approach proposed here, enables more intuitive and interpretable results by preserving the system-wide population and level of service. In this way, a disparity index is useful to highlight the absolute difference in accessible provider-to-population ratios across population units.

The disparity for SoBi is shown in the following map. The figure reveals the spatial distribution in disparity, with levels of access that are lower than the mean in the periphery of the service area, compared to levels of access that are greater than the mean in the higher-density and more connected urban center.

```{r}
population_sobi %>%
  mutate(Lei = accessibility - sum(sobi_los$sobi_los)/nrow(population_sobi %>% drop_na(accessibility))) %>%
  ggplot() +
  geom_sf(aes(color = Lei)) +
  geom_sf(data = sobi_service,
          fill = NA) +
  scale_color_gradient2()
```

```{r accessibility_disparities, echo=FALSE, include=FALSE}
#Disparities in this case is defined as the difference between the calculated accessibility (which is interpreted as the local provider-to-population ratio) and the local provider-to-population ratio if the level of service was distributed uniformly
accessibility_disparities <- transmute(accessibility, DA, 
                                      Binary = Binary_Adjusted - sum(los_binary_adjusted)/nrow(accessibility_binary_adjusted),
                                      Stepwise = Stepwise_Adjusted- sum(los_stepwise_adjusted)/nrow(accessibility_stepwise_adjusted))
```

```{r join-geography, echo=FALSE, include=FALSE}
#Join accessibility measures to geography.
hamilton_da.sf_accessibility <- left_join(hamilton_da.sf, accessibility, by = "DA")
hamilton_da.sf_comparison <- left_join(hamilton_da.sf, accessibility_comparison, by = "DA")
hamilton_da.sf_disparities <- left_join(hamilton_da.sf, accessibility_disparities, by = "DA")
```
